---
title: "CCA Results"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
library(ggplot2)
library(candisc)
knitr::opts_chunk$set(echo = TRUE)

dens_plot <- function(null_ccs, Z_table, xlim = c(0,1), label_h = 9, text_offset = .075, nudge_x = -.025){
  lt <- 'solid'
  segalpha <- 1
  segcol <- '#bb6a23'
  dotfill <- '#ffd653'
  denscol <- '#2a5078'
  densfill <- '#a7c6b8'
  p <- ggplot(null_ccs, aes(x = cc)) +
    geom_density(fill = densfill, color = denscol, alpha = 1) + 
    geom_segment(data = Z_table, aes(x = cc_obs, xend = cc_obs, y = 0, yend = label_h), color = segcol,
                 linetype = lt, alpha = segalpha) +
    geom_segment(data = Z_table, aes(x = cc_obs, xend = cc_obs - text_offset, y = label_h, yend = label_h), color = segcol,
                 linetype = lt, alpha = segalpha) +
    geom_point(data = Z_table, aes(x = cc_obs, y = 0), color = segcol, fill = dotfill, shape = 21, size = 2.5) + 
    # geom_point(data = Z_table, aes(x = cc_obs - text_offset, y = label_h), color = '#2a5078', shape = 18) + 
    geom_text(data = Z_table, aes(label = sprintf('Z = %.2f', Z), x = cc_obs - text_offset, y = label_h), 
              hjust = 1, vjust = .5, nudge_x = nudge_x, nudge_y = 0, size = 3,
              color = '#444444',
              family = 'NimbusSanCond') + 
    theme_minimal() + 
    scale_x_continuous(breaks = sort(c(xlim, min(xlim) + diff(xlim)/2))) + 
    coord_cartesian(x = xlim) + 
    facet_wrap(~ k, ncol = 5) + 
    labs(x = 'Canonical correlation across 10,000 permutations', y = 'Canonical variate number')
  return(p)
}

```

# Drysdale selection method

## Load permutations

```{r}
perms_list <- lapply(dir('data', pattern = 'cca_perms-drysdale-chunk_.*rds', full.names = TRUE), readRDS)

#The optimal penalty for each permuted data set
penalties <- unlist(lapply(perms_list, function(permset){
  lapply(permset, function(aperm) unique(aperm$penalty))
}))
#Create a data frame with one row for each correlation for each permutation.
#Column `k` indexes the correlation.
null_ccs <- data.table::rbindlist(
  lapply(perms_list, function(permset){
    data.table::rbindlist(
      lapply(
        permset, 
        function(aperm) data.frame(cc = aperm$cca_cors, k = 1:length(aperm$cca_cors))))
  }))
#index the permutation iteration too, just in case we want to see what penalty was used.
null_ccs[, i := 1:.N, by = k]
null_ccs[, mean := mean(cc), by = k]
```

## Load CCA of observed data

```{r}
cca_obsv <- readRDS(file.path('data', 'cca-drysdale.rds'))
cca_penalty_used <- cca_obsv$cca.permute_obj$bestpenaltyx
cc <- data.frame(cc_obs = cca_obsv$cca_obj$cors, 
                 k = 1:length(cca_obsv$cca_obj$cors))
#To find the Z score we just see what proportion of null CCs are less than the
#observed CC and then use qnorm to turn that probability into a Z score. We can
#then get a two-sided p-value for the CC using pnorm (there are other ways
#too--we could just directly use the prop_null_lt_obs).
Z_table <- null_ccs[cc, on = 'k'][, list(prop_null_lt_obs = mean(cc < cc_obs),
                                         Z = qnorm(mean(cc < cc_obs)),
                                         `P(>|Z|)` = pnorm(abs(qnorm(mean(cc < cc_obs))), lower.tail = F)*2,
                                         cc_obs = unique(cc_obs)), by = k]
```

```{r, fig.width=5, fig.height=3, fig.cap='Distribution of penalties across permutations. Blue line indicates penalty used for unpermuted data'}
ggplot(data.frame(penalties = penalties), aes(x = penalties)) + 
  geom_histogram(bins = 19, fill = '#a7c6b8') + 
  geom_vline(xintercept = cca_penalty_used, color = '#2a5078') + 
  theme_minimal()
```

```{r, fig.width=10, fig.height=4,fig.cap='Density of canonical correlations (CC) on permuted data for each of 10 canonical variates in order. Lines indicate CC of unpermuted data with Z scores for their location in the null density.'}
#https://www.instagram.com/p/B_fRZQzgsnc/
dens_plot(null_ccs = null_ccs, Z_table = Z_table)
```

# Drysdale 2nd selection method

## Load permutations

```{r}
perms_list <- lapply(dir('data', pattern = 'cca_perms-drysdale2-chunk_.*rds', full.names = TRUE), readRDS)

#The optimal penalty for each permuted data set
penalties <- unlist(lapply(perms_list, function(permset){
  lapply(permset, function(aperm) unique(aperm$penalty))
}))
#Create a data frame with one row for each correlation for each permutation.
#Column `k` indexes the correlation.
null_ccs <- data.table::rbindlist(
  lapply(perms_list, function(permset){
    data.table::rbindlist(
      lapply(
        permset, 
        function(aperm) data.frame(cc = aperm$cca_cors, k = 1:length(aperm$cca_cors))))
  }))
#index the permutation iteration too, just in case we want to see what penalty was used.
null_ccs[, i := 1:.N, by = k]
null_ccs[, mean := mean(cc), by = k]
```

## Load CCA of observed data

```{r}
cca_obsv <- readRDS(file.path('data', 'cca-drysdale2.rds'))
cca_penalty_used <- cca_obsv$cca.permute_obj$bestpenaltyx
cc <- data.frame(cc_obs = cca_obsv$cca_obj$cors, 
                 k = 1:length(cca_obsv$cca_obj$cors))
#To find the Z score we just see what proportion of null CCs are less than the
#observed CC and then use qnorm to turn that probability into a Z score. We can
#then get a two-sided p-value for the CC using pnorm (there are other ways
#too--we could just directly use the prop_null_lt_obs).
Z_table <- null_ccs[cc, on = 'k'][, list(prop_null_lt_obs = mean(cc < cc_obs),
                                         Z = qnorm(mean(cc < cc_obs)),
                                         `P(>|Z|)` = pnorm(abs(qnorm(mean(cc < cc_obs))), lower.tail = F)*2,
                                         cc_obs = unique(cc_obs)), by = k]
```

```{r, fig.width=5, fig.height=3, fig.cap='Distribution of penalties across permutations. Blue line indicates penalty used for unpermuted data'}
ggplot(data.frame(penalties = penalties), aes(x = penalties)) + 
  geom_histogram(bins = 19, fill = '#a7c6b8') + 
  geom_vline(xintercept = cca_penalty_used, color = '#2a5078') + 
  theme_minimal()
```

```{r, fig.width=10, fig.height=4,fig.cap='Density of canonical correlations (CC) on permuted data for each of 10 canonical variates in order. Lines indicate CC of unpermuted data with Z scores for their location in the null density.'}
#https://www.instagram.com/p/B_fRZQzgsnc/
dens_plot(null_ccs = null_ccs, Z_table = Z_table)
```

# Drysdale 3nd selection method

This is the closest to an exact replication of Drysdale et al's method (as reported by Dinga et al). We select 100 features from the resting state data that show the highest absolute maximum correlations with any of the threat/deprivation features, and 100 features from the threat/dep data that show the highest absolute max correlation with any of the resting state features. We then submit these features to a non-regularized CCA.

## Load permutations

```{r}
perms_list <- lapply(dir('data', pattern = 'cca_perms-drysdale3-noreg--chunk_.*rds', full.names = TRUE), readRDS)

#Create a data frame with one row for each correlation for each permutation.
#Column `k` indexes the correlation.
null_ccs <- data.table::rbindlist(
  lapply(perms_list, function(permset){
    data.table::rbindlist(
      lapply(
        permset, 
        function(aperm) data.frame(cc = aperm$cca_cors, k = 1:length(aperm$cca_cors))))
  }))
#index the permutation iteration too, just in case we want to see what penalty was used.
null_ccs[, i := 1:.N, by = k]
null_ccs[, mean := mean(cc), by = k]
```

## Load CCA of observed data

```{r}
cca_obsv <- readRDS(file.path('data', 'cca-drysdale3-noreg-.rds'))
cc <- data.frame(cc_obs = cca_obsv$cca_obj$cancor, 
                 k = 1:length(cca_obsv$cca_obj$cancor))
#To find the Z score we just see what proportion of null CCs are less than the
#observed CC and then use qnorm to turn that probability into a Z score. We can
#then get a two-sided p-value for the CC using pnorm (there are other ways
#too--we could just directly use the prop_null_lt_obs).
Z_table <- null_ccs[cc, on = 'k'][, list(prop_null_lt_obs = mean(cc < cc_obs),
                                         Z = qnorm(mean(cc < cc_obs)),
                                         `P(>|Z|)` = pnorm(abs(qnorm(mean(cc < cc_obs))), lower.tail = F)*2,
                                         cc_obs = unique(cc_obs)), by = k]
```

```{r, fig.width=10, fig.height=4,fig.cap='Density of canonical correlations (CC) on permuted data for each of 10 canonical variates in order. Lines indicate CC of unpermuted data with Z scores for their location in the null density. X axis starts at r=.90.'}
#https://www.instagram.com/p/B_fRZQzgsnc/
dens_plot(null_ccs = null_ccs[k <= 10], Z_table = Z_table[k <= 10], xlim = c(.9, 1), label_h = 50, text_offset = .02, nudge_x = -.0025)
```

```{r}
library(data.table)
w_dt <- data.table::data.table(Wilks(cca_obsv$cca_obj))
Z_table_table <- copy(Z_table[, c('Z', 'P(>|Z|)', 'cc_obs')])
setnames(Z_table_table, c('cc_obs', 'Z'), c('CanR', 'Z_perm'))
knitr::kable(w_dt[Z_table_table, on = 'CanR'], caption = 'Wilk\'s test (F) and permutation Z score for observed canonical correlation')
```

# Full Selection via regularization

No pre-selection of features. 

## Load permutations

```{r}
perms_list <- lapply(dir('data', pattern = 'cca_perms-nofeatsel-chunk_.*rds', full.names = TRUE), readRDS)

#The optimal penalty for each permuted data set
penalties <- unlist(lapply(perms_list, function(permset){
  lapply(permset, function(aperm) unique(aperm$penalty))
}))
#Create a data frame with one row for each correlation for each permutation.
#Column `k` indexes the correlation.
null_ccs <- data.table::rbindlist(
  lapply(perms_list, function(permset){
    data.table::rbindlist(
      lapply(
        permset, 
        function(aperm) data.frame(cc = aperm$cca_cors, k = 1:length(aperm$cca_cors))))
  }))
#index the permutation iteration too, just in case we want to see what penalty was used.
null_ccs[, i := 1:.N, by = k]
null_ccs[, mean := mean(cc), by = k]
```

## Load CCA of observed data

```{r}
cca_obsv <- readRDS(file.path('data', 'cca-nofeatsel.rds'))
cca_penalty_used <- cca_obsv$cca.permute_obj$bestpenaltyx
cc <- data.frame(cc_obs = cca_obsv$cca_obj$cors, 
                 k = 1:length(cca_obsv$cca_obj$cors))
#To find the Z score we just see what proportion of null CCs are less than the
#observed CC and then use qnorm to turn that probability into a Z score. We can
#then get a two-sided p-value for the CC using pnorm (there are other ways
#too--we could just directly use the prop_null_lt_obs).
Z_table <- null_ccs[cc, on = 'k'][, list(prop_null_lt_obs = mean(cc < cc_obs),
                                         Z = qnorm(mean(cc < cc_obs)),
                                         `P(>|Z|)` = pnorm(abs(qnorm(mean(cc < cc_obs))), lower.tail = F)*2,
                                         cc_obs = unique(cc_obs)), by = k]
```

```{r, fig.width=5, fig.height=3, fig.cap='Distribution of penalties across permutations. Blue line indicates penalty used for unpermuted data'}
ggplot(data.frame(penalties = penalties), aes(x = penalties)) + 
  geom_histogram(bins = 19, fill = '#a7c6b8') + 
  geom_vline(xintercept = cca_penalty_used, color = '#2a5078') + 
  theme_minimal()
```

```{r, fig.width=10, fig.height=4,fig.cap='Density of canonical correlations (CC) on permuted data for each of 10 canonical variates in order. Lines indicate CC of unpermuted data with Z scores for their location in the null density.'}
#https://www.instagram.com/p/B_fRZQzgsnc/
dens_plot(null_ccs = null_ccs[k <= 10], Z_table = Z_table[k <=10])
```


# Xia selection method

## Load permutations

```{r}
perms_list <- lapply(dir('data', pattern = 'cca_perms-xia.*rds', full.names = TRUE), readRDS)

#The optimal penalty for each permuted data set
penalties <- unlist(lapply(perms_list, function(permset){
  lapply(permset, function(aperm) unique(aperm$penalty))
}))
#Create a data frame with one row for each correlation for each permutation.
#Column `k` indexes the correlation.
null_ccs <- data.table::rbindlist(
  lapply(perms_list, function(permset){
    data.table::rbindlist(
      lapply(
        permset, 
        function(aperm) data.frame(cc = aperm$cca_cors, k = 1:length(aperm$cca_cors))))
  }))
#index the permutation iteration too, just in case we want to see what penalty was used.
null_ccs[, i := 1:.N, by = k]
null_ccs[, mean := mean(cc), by = k]
```

## Load CCA of observed data

```{r}
cca_obsv <- readRDS(file.path('data', 'cca-xia.rds'))
cca_penalty_used <- cca_obsv$cca.permute_obj$bestpenaltyx
cc <- data.frame(cc_obs = cca_obsv$cca_obj$cors, 
                 k = 1:length(cca_obsv$cca_obj$cors))
#To find the Z score we just see what proportion of null CCs are less than the
#observed CC and then use qnorm to turn that probability into a Z score. We can
#then get a two-sided p-value for the CC using pnorm (there are other ways
#too--we could just directly use the prop_null_lt_obs).
Z_table <- null_ccs[cc, on = 'k'][, list(prop_null_lt_obs = mean(cc < cc_obs),
                                         Z = qnorm(mean(cc < cc_obs)),
                                         `P(>|Z|)` = pnorm(abs(qnorm(mean(cc < cc_obs))), lower.tail = F)*2,
                                         cc_obs = unique(cc_obs)), by = k]
```

```{r, fig.width=5, fig.height=3, fig.cap='Distribution of penalties across permutations. Blue line indicates penalty used for unpermuted data'}
ggplot(data.frame(penalties = penalties), aes(x = penalties)) + 
  geom_histogram(bins = 19, fill = '#a7c6b8') + 
  geom_vline(xintercept = cca_penalty_used, color = '#2a5078') + 
  theme_minimal()
```

```{r, fig.width=10, fig.height=4,fig.cap='Density of canonical correlations (CC) on permuted data for each of 10 canonical variates in order. Lines indicate CC of unpermuted data with Z scores for their location in the null density.'}
dens_plot(null_ccs = null_ccs, Z_table = Z_table)
```